\documentclass[preview]{standalone}
%\usepackage{prelude}
\input{prelude}

\begin{document}
\section{Outlook}

In this bachelor thesis we proposed a concept to preprocess data for visualization, which we called view. We presented, defined and discussed types and operations on views and proposed example views. For some of them, use cases were presented to show their applicability in practice. Apart from the formalization, all features of views of views and all proposed views have been implemented in the context of the \pmcvis project. The implementation is based on a fast but memory hungry graph structure provided by the \jgrapht library.

In conclusion, the formalization concisely describes the notion of a concept view, which helps to understand the implementation and may be useful for the development of other simplification approaches in model checking related systems. For the proposed example views, it should be noted that the author had little experience in working with \mdpsN. Therefore, most of the views are not motivated by how a view might be useful when working with \mdpN, but rather what simplifications are possible on \mdpsN, taking into account its components and structure. The use cases shown are only a small selection of what might be possible. The real practicability will be proven in practice. Also, \mdpN-experienced users who actually work with views in practice will most likely raise other ideas for views that might be helpful.

In addition to practical advancements for views and their use cases, there is still room for concept exploration from a theoretical perspective. For one, considering views with probabilities would have been beyond the scope of a bachelor thesis, so the probabilities present in \mdpsN were not explored. Furthermore, there may be advanced concepts from other disciplines that can be used to gain simplifications on \mdpsN.

The provided implementation serves as a solid framework to implement further views. As mentioned before, it is quite memory intensive. The \mdpN used for testing with 5 million states could not be built on the test machine due to lack of memory. Thus, for a given amount of available memory, the data structure imposes a hard limit on the applicability of views in practice if the \mdpN exceeds a certain size. This problem can be solved by not using the data structure (\texttt{mdpGraph}), but by accessing the database directly. This allows views to be used on almost arbitrarily large \mdpsN. While this is a great advantage, its downside is an immense performance loss (increase in build times) compared to the used \texttt{mdpGraph}, because database accesses with the current version of \pmcvis are orders of magnitude slower than accessing the information in the \texttt{mdpGraph}. This is due to the fact that the \texttt{mdpGraph} consists of very efficient data structures such as hash maps and hash sets. In addition, the \texttt{mdpGraph} and its data structures are stored in memory, while accessing the database means accessing the machine's memory, which is in principle much slower than accessing memory. This leads to immense execution times when creating a view, especially if stateful database accesses are to be performed.


\end{document}