\documentclass[preview]{standalone}
%\usepackage{prelude}
\input{prelude}

\begin{document}
\section{Introduction}

\begin{itemize}
	\item PREWRITTEN INTRODUCTION
	\item modern world heavily relies on ICT (Information and Communication Technology) systems
	\item \redcomment{they are found in} devices used everyday as smartphones or laptops or in distributed systems such as the infrastructure sustaining the internet, but also in live saving ones utilized in medicine.
	\item apart from performance and provided features and functionality, one of the most relevant aspects of such systems is their faulty free behavior, when active or running.
	\item Although desirable many people have personal experience with such systems not behaving as expected, what is commonly referred to as a bug.
	\item These can reach from a small disturbance in the user experience to inoperability of the whole system. Depending on the system the impact could be an annoyed user or financial damage of several million.
	\item To prevent these possibly severe negative effects, methods to verify the correct behavior of a system are needed.
	\item there exist several verification techniques
	\item One approach for exhaustive is Model checking - a formal model-based method for system verification.
	\item as the name suggests it is based on models describing the possible system behavior in a mathematically precise and unambigous manner. A state in the model refers to a possible state of the system. In precise terms, model checking is an automated technique, to check if a given property holds for every state of a given finite-state model. \cite[chs. 1.1 and 1.2]{Baier2008}.
	\item Probabilistic model checking allows not only nondeterministic transitions in the models checked, but also probabilistic ones. It enables to properly model and check systems in which also stochastic phenomena are occurring. 
	\item The major limitation for algorithms running on the set of states of models, in order to model check them, is the state explosion problem. The state explosion problem states that the number of states grow exponentially in the number of variables in a program graph or the number of components in concurrent systems \cite[ch. 2.3]{Baier2008}. Already simple system models can lead to complex system behavior and an immense amount of states. This is not only a problem for algorithms but also humans who need to understand, analyze and review models in the context of model checking.
	\item An interactive visualization can assist with this issue and even be made use of to achieve further goals such as debugging or model repair. However, the pure data volume is left unaffected.
	\item There are visualization techniques for multivariate networks (networks where nodes and relationships have attributes) which can be used for larger amounts of data. But even these perform poorly for very large amounts of data as they are easily achievable with models used in model checking \cite{Kerren2014,Nobre2019}. 
	\item There is graph abstraction (in Visualization) but we wont look further into that
	\item Moreover they represent approaches that are not specific to the domain.
	\item In this thesis we want to explore approaches that use domain specific information and knowledge to preprocess data, that is then utilized for visualization.
	\item Preprocessing is an approach to clean up and prepare data for Data Mining and Machine Learning. The goal is to prepare and clean up data so that they meet the requirements in order for algorithms being able to analyze them. In order to achieve that data is reduced if there is too little, inferred if there is not sufficient or cleaned if it includes noise or errors. Since the state explosion problem causes an immense amount of states approaches that reduce the amount of data are of interest. Methods reducing the amount of data either reduce the size of a single sample (feature selection), reduce the amount of samples (data elimination) or cluster samples to one single sample (data selection)\cite{Famili1997,Garcia2016}. Even though such methods exist they are not domain specific. Still concepts presented in this thesis are similar to the notion data selection.
	Apart from
	
	\item Apart from \redcomment{XXXX} there exist domain specific to simplify \mdpsN. 
	\item 
	\item These define equivalence or order relations on the set of states, if these states can simulate each other stepwisely or in several steps with respect to atomic propositions, which causes preservation of certain logical formulae used in model checking.
	\item Although certain logical formulae and hence model checking properties are preserved other information is lost. In addition the computation of these computations is rather costly.
	\item We introduce a concept, which will be called \viewN, that is very similar to abstraction \cite[pp. 499]{Baier2008}, but without permanently losing on information. Moreover, its intention is to show humans interacting with the visualization as much information possible in compact and concise form, rather than preserving logical formulae.
	\item After giving some fundamentals about the systems where the concept \viewN may be applied in chapter \ref{ch:prelim} we will formalize \viewsN, discuss types and operations on and with them in chapter \ref{ch:view}.
	\item  In chapter \ref{ch:viewexmp} we will give examples of views utilizing \chgphN components, \viewsN utilizing the \chgphN structure and \viewsN utilizing results of model checking.
	\item For this thesis the scope of views considered is limited to those that do not take advantage of the probabilities in \chgphsN.
	\item In chapter \ref{ch:viewimpl} we will elaborate on where and how the proposed \viewsN of chapter \ref{ch:viewexmp} have been implemented
	\item Lastly the \viewsN will be evaluated in chapter \ref{ch:eval} by considering three usecases how \viewsN can be applied and used. Moreover there will be an overview about performance and scalability of the proposed \viewsN. 
	
\end{itemize}

\begin{itemize}
	\item INTRODUCTION STRUCTURE (used above)
	\item Objective of this thesis: Preprocessing of pmc-visualization
	\item where pmc probablistic model checking
	\item model checking: formal model-based method for system verification that is based on models describing the possible system behavior in mathematically precise and unambigous manner. A state in the model is refers to a possible state of the system. More precisely model checking is an automated technique for given finite-state model and given property, to check if the property holds for every state  \cite[chs. 1.1 and 1.2]{Baier2008}
	\item model checking
	\item problems with model checking
	\item probabilistic model checking allows not only nondeterministic transitions in the models checked, but also probabilistic ones. It is enabling to properly model and check systems in which also occur stochastic phenomena. 
	\item The major limitation for algorithms running on the set of states of models, in order to model check them, is the state explosion problem. The state explosion problem states that the number of states grow exponentially in the number of variables in a programm graph or the number of components in concurrent systems \cite[ch. 2.3]{Baier2008}. Already simple system models can lead to complex system behavior and an immense amount of states. This is not only a problem for algorithms but also humans who need to understand, analyze and review models in the context of model checking.
	\item state exponential growth of states in number of variables in a programm graph or the number of components in concurrent system \cite[ch. 2.3]{Baier2008}
	\item interactive visualization can help
	\item visualization has some tools of its own as in 
	\item not domain specific 
	\item want to use domain specific to preprocess data for visualization
	\item preprocessing is ... 
	\item includes teqnuiqes as
	\item normally used for algos -> some of general approaches usable but concrete techniques not really
	\item graph abstractions
	\item from pmc book
	\item what we want to is very similar
	\item these have disadvantge of losing information
	\item after giving some fundamentals about the systems where views may be applied in Chapter 2 we will the concept of
	\item formalize views, discuss types of views and operations on and with them 
	\item  In chaptergive examples of views utilizing mdp components, view utilizing the mdp structure and utilizing results of model checking
	\item limit scope to views that do not take adavantage of the probabilities of mdps
	\item evaluate them by considering three usecases how views can be applied and used in order to help understanding the graph
	\item thesis is aimed introducing a concept to get simplified representations on mdps and exploring ideas for such simplifcations.
	\item it is written in the context of a project at the chair developing a predicitive model checking visualization tool in which some of the ideas may be used.
	
\end{itemize}

\end{document}
